---
title: "Problem with median for South Africa data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First check the median bin

```{r}
library(binsmooth)
binmax <-  c(1, 4800, 9600, 19200, 38400, 76800, 153600, 307200, 614400, 1228800, 2457600, NA) 
bincounts <- c(663, 142, 209, 673, 1047, 977, 685, 287,  67, 3, 5, 1) #The number of household for each  bin
median_bin <- which(cumsum(bincounts) > sum(bincounts)/2)[1]
cat("Median should be between", binmax[median_bin-1], "and", binmax[median_bin], "\n")
```

## Median is way off

```{r}
splb <- splinebins(binmax, bincounts)
median_income <- sb_percentiles(splb, p = 50) #Get median
median_income
```

Things are slightly better if you change the spline method, but not much.

```{r}
splb <- splinebins(binmax, bincounts, monoMethod = "monoH.FC")
median_income <- sb_percentiles(splb, p = 50) #Get median
median_income
```

The upper bound of the tail is clearly nonsense.

```{r}
splb <- splinebins(binmax, bincounts)
cat("Upper bound of support of PDF", splb$E, "\n")
plot(splb$splinePDF, 0, splb$E)
```

Since the bin width doubles with each bin, I wonder if taking the log would make things behave better. Nope.

```{r}
lbinmax <- log(binmax)
lsplb <- splinebins(lbinmax, bincounts)
lmedian_income <- sb_percentiles(lsplb, p = 50)
cat("Estimated median:", exp(lmedian_income))
```

Instead of making the first bin start at 1, let's just put the zero income people into the second bin and not use the first bin.

```{r}
binmax <-  c(4800, 9600, 19200, 38400, 76800, 153600, 307200, 614400, 1228800, 2457600, NA) 
bincounts <- c(663+142, 209, 673, 1047, 977, 685, 287,  67, 3, 5, 1) #The number of household for each  bin
median_bin <- which(cumsum(bincounts) > sum(bincounts)/2)[1]
cat("Median should be between", binmax[median_bin-1], "and", binmax[median_bin], "\n")
splb <- splinebins(binmax, bincounts)
median_income <- sb_percentiles(splb, p = 50) #Get median
cat("Splined median:", median_income, "\n")
lbinmax <- log(binmax)
lsplb <- splinebins(lbinmax, bincounts)
lmedian_income <- sb_percentiles(lsplb, p = 50)
cat("Estimated median using logs:", exp(lmedian_income-1))
```

```{r}
library(pracma)
binmax <-  c(1, 4800, 9600, 19200, 38400, 76800, 153600, 307200, 614400, 1228800, 2457600, NA) 
bincounts <- c(663, 142, 209, 673, 1047, 977, 685, 287,  67, 3, 5, 1) #The number of household for each  bin
sb <- stepbins(binmax, bincounts)
plot(sb$stepPDF, do.points = FALSE)
integral(sb$stepPDF, 0, sb$E) # should be approximately 1
integral(function(x){1-sb$stepCDF(x)}, 0, sb$E) # should be the mean
```

Let's see if these bin counts work for more uniform bins.

```{r}
binmax <-  c(10000,15000,20000,25000,30000,35000,40000,45000,50000,60000,NA)
bincounts <- c(663+142, 209, 673, 1047, 977, 685, 287,  67, 3, 5, 1) #The number of household for each  bin
median_bin <- which(cumsum(bincounts) > sum(bincounts)/2)[1]
cat("Median should be between", binmax[median_bin-1], "and", binmax[median_bin], "\n")
splb <- splinebins(binmax, bincounts)
median_income <- sb_percentiles(splb, p = 50) #Get median
cat("Splined median:", median_income, "\n")
plot(splb$splinePDF, 0, 100000)
```


